Heritability

### 2021-08-17 Manuscript/manuscript.Rmd
update manuscript and analyses and figures plus add citations

---
title: Artificial ecosystem selection to deduce the mapping between microbial community structure and ecosystem function
author: "Andrew H. Morris and Brendan J. M. Bohannan"
output:
  word_document:
    reference_docx: reference.docx
editor_options:
  chunk_output_type: inline
---

@@

```{r initialize_packages}
library(tidyverse)
library(captioner)
library(broom)
library(broman)
library(morris)
#library(metafor)
```
```{r load_data}
heritability <- read_tsv('../Output/heritability.tsv')
fluxes <- read_tsv('../Output/fluxes.tsv')

@@

figures <- captioner()
figures("response", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment. Deviance is calculated as the mean of P within each passage minus the mean of N within each passage. Errorbars are the standard error of the difference between P and N for each passage. The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations. The gray ribbon is the 95% confidence interval for the regression.")
figures("herit", "Regression of offspring CH~4~ flux on mid-parent CH~4~ flux. Mid-parent is the mean of jars selected to inoculate the next passage. The mid-offsprings are the mean of all twelve jars produced in one passage. The best fit line is an ordinary least squares regression line and the gray ribbon is the 95% confidence interval for the regression coefficient.")

@@

One of the fundamental questions in Ecology is To what extent does biodiversity
drive variation in ecosystem function? Variation in abiotic conditions
regulates the rate of ecosystem processes, but it remains unclear to what
extent variation in community structure regulates the rate of ecosystem
processes. Many ecosystem functions are mediated by organisms. In particular,
microorganisms such as bacteria mediate the major biogeochemical cycles such
as the carbon and nitrogen cycles. Despite the central role of bacteria in
performing these processes, we have not found a consistent relationship between
microbial community structure and the rate of an ecosystem process. In order
to understand the importance of biodiversity to ecosystem functioning and to
better predict and manage specific ecosystem processes, we need to determine
the relationship between microbial community structure and ecosystem function.

In macroorganismal communities, there appears to be a relationship between the
diversity of a community and the resilience of ecosystem functions to
environmental perturbations (Tilman, ????... Hooper). Typically, this is the
diversity (or more specifically, the richness) of a particular functional
group, such as photosynthetic plants. Similar studies in microbial communities
rarely find a relationship between diversity and function (Graham, ). However,
these studies typically define diversity much more broadly than plant diversity
studies by looking at the phylogenetic diversity of the entire bacterial and
archaeal community. Studies that focus more narrowly on a particular ecosystem
process and a functional group involved in that process still rarely find a
significant correlation between the abundance of a functional group and the
rate of the corresponding process (Rocca, ????). For certain ecosystem
functions in certain ecosystems, these relationships can be informative. For
example, the abundance of a single methanogenic species (*G. species*) is
correlated with methane emissions in a permafrost ecosystem (cite). In
addition, bacterial and archaeal community composition is correlated with
methane emissions in forests and pastures of the Brazilian Amazon (Meyer et al.
2020).  Therefore, there appears to be some indication that microbial community
structure drives variation in methane cycling in soils.

To further investigate the relationship between microbial community structure
and the rate of methane cycling in soils, we want to determine which organisms
are important for regulating methane cycling. This ecological question is
analogous to the problem of genotype-phenotype mapping in quantitative genetics
(Morris 2020). Variation in an organismal trait is partially determined by
environmental conditions and partially determined by genetic variation. One goal
of quantitative genetics is to determine the proportion of the variation
attributable to genetics and to identify which alleles regulate the phenotypic
trait. These goals are commonly achieved through artificial selection
experiments and association mapping studies, such as genome-wide association
studies. To determine the mapping between microbial community structure and
ecosystem function, we applied these approaches of genotype-phenotype mapping
to whole microbial communities. In this case, the "alleles" are different
microbial taxa and the "phenotype" is the rate of ecosystem function at the
whole-ecosystem level.

@@

## Selection experiment
We performed an artificial ecosystem selection experiment similar to Swenson et
al. (2000) by passaging replicate soil microbial communities. The trait we
selected on was CH~4~ oxidation rate. Our experiment had two selection lines
with twelve jars each for a total of twenty-four jars per passage. One line was
a positive selection line where the two or three jars with the highest CH~4~
oxidation rate were chosen to inoculate the next set of positive jars. The
other line was a neutral selection line where an equal number of jars as the
positive line were chosen at random to inoculate the next set of neutral jars.
The number of jars chosen was based on the distribution of fluxes among the
positive jars, i.e., we chose the top three jars unless only two jars had
considerably greater oxidation rates based on visual inspection of histograms.
The experiment was carried out over five passages.

The initial soil microbial community was sampled from the top 10 cm of an
upland mineral soil under a deciduous forest ecosystem near the University of
Oregon campus in Eugene, OR, USA. Incubations were performed in 500 mL mason
jars (Ball Corporation, Broomfield, Colorado, USA) with rubber septa installed
in the lids. Each jar was sterilized with 70% ethanol to which was added 45 g
of autoclaved potting mix, 5 g of living soil, and 3.5 mL of sterile deionized
water to bring them to <mark> 60%??? of field capacity</mark>. The jars were
then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
headspace concentration to approximately 1000 ppm CH~4~. To create the two
treatment groups, twenty-four jars were created in an identical manner and then
randomly assigned to either the positive or neutral selection line. Jars were
flushed and respiked twice per week to maintain elevated CH~4~ concentrations
and were incubated at ambient temperature for approximately three weeks.
Methane oxidation rates were determined at the end of the incubation period.
These values were then used to determine the jars selected to inoculate the
next set of jars.

@@

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a
<mark>???</mark> gas chromatograph fitted with an electron capture detector
(company, city, state, USA) to determine the headspace CH~4~ concentration.
Fluxes were calculated from a first-order exponential decay function as decay
constant `k` with units day^-1^. Oxidation rates are presented as the additive
inverse of `k` (i.e., `-k`) so that a more positive value represents a faster
oxidation rate. The jars selected for the positive treatment in passage 2 had
the lowest methane oxidation rate of the twelve jars due to a calculation error
in the methane oxidation rate. All other passages correctly used the top three jars.

@@

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at -80$^\circ$C for later DNA extraction.
Soil DNA was extracted from a 0.25 g subsample of soil. Negative controls were extracted
from autoclaved potting mix and DNase-free water. Extractions were performed
using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany) and quantified using
Qubit (company, city, state). In order to estimate the diversity and relative
abundance of the bacterial and archaeal taxa in our soil ecosystems, we
sequenced the V3-V4 region of the 16S rRNA gene using the Earth Microbiome
Project protocol and primers<mark>???</mark>. Sequencing was performed on the
Illumina NovaSeq 6000 with paired-end 150 bp reads.

@@

Bioinformatic analyses were conducted using the `QIIME2`
bioinformatics platform (Bolyen et al. 2019). Raw sequencing reads were
demultiplexed using ??? (citation), denoised using `DADA2` (citation).
Taxonomic assignment for the 16S reads was performed using ??? based on the
Silva database. All subsequent tatistical analyses were performed in the `R`
statistical programming environment (R Core Team 2020). Richness estimates were
made using the `breakaway` package and diversity indices using the `DivNet`
package (Willis et al. ???).

@@

To test whether there was a significant change in CH~4~ oxidation rate (k) as a
response to selection, we tested a difference in slopes between the positive
and neutral selection lines. The CH~4~ oxidation rates were strongly right
skewed with most values close to zero and few large, positive values. This
resulted in residuals that did not meet the assumptions of constant variance
and normal distribution. Therefore, methane oxidation rates were log~10~
transformed to better meet the assumptions of a linear model and to make
figures easier to interpret. First, we tested the effect of treatment by
fitting two nested models with and without treatment using `lm` in `R`. The
model formulation for the full model was `flux ~ passage * treatment` and the
model formulation for the reduced model was `flux ~ passage`. We compared these
models using the likelihood ratio test with the `anova` function in `R`. We
then fit the full model to determine the slope of the positive line, which
represents the change in methane oxidation rate per passage as a response to
selection.

@@

To estimate the proportion of variance in ecosystem function due to additive
genetic variance of the microbial communtiy, we applied the Breeder's equation
to the results of our selection experiment. the heritability of
CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense
heritability ($h^2$) as the regression of offspring on mid-parent (Goodnight
2000, Falconer 1997). The mid-parent was the mean for all three selected jars
and the offspring was the mean for all jars produced by those parents.

In addition, we calculated the inbreeding coefficient, $F$, as:

$$
F = 1 - \left(1 - \frac{1}{2N} \right) ^{T}
$$

where $N$ is the population size and $T$ is the number of generations (Goodnight 2000 or Falconer 1997).

### Taxa differentially abundant in the positive selection treatment

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance using a beta-binomial model (Martin et al. 2020). We
compared taxa in passage 5 between the positive and neutral treatment to
identify taxa that

@@

@@

@@

We observed a response to selection on soil methane oxidation rate at the whole
ecosystem level (`r figures("response", display="cite")`).
At the start of the experiment, the Positive
treatment had a mean methane oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`%
lower than the
Neutral treatment (SE = 
`r myround(tidy(response_model)[[3, 3]], 2)`, t = 
`r myround(tidy(response_model)[[3, 4]], 2)`, p = 
`r myround(tidy(response_model)[[3, 5]], 2)`). There was no change in methane oxidation rate of the Neutral
treatment over the five passages (slope = 
`r myround(tidy(response_model)[[2, 2]], 2)`, SE = 
`r myround(tidy(response_model)[[2, 3]], 2)`, t = 
`r myround(tidy(response_model)[[2, 4]], 2)`, p = 
`r myround(tidy(response_model)[[2, 5]], 2)`). By contrast, the Positive treatment had a
`r round(per_slope)`% increase in methane
oxidation rate per passage (slope = 
`r myround(tidy(response_model)[[4, 2]], 2)`, SE = 
`r myround(tidy(response_model)[[4, 3]], 2)`, t =
`r myround(tidy(response_model)[[4, 4]], 2)`, p = 
`r myround(tidy(response_model)[[4, 5]], 2)`).

@@

Here, I fit two models. The first is simply offspring flux on selected parent
flux, ignoring treatment. There is no correlation. Next, I added treatment and the interaction between treatment and parental
flux. Here, there is a significant correlation between parent and offspring for
both the Neutral treatment `r print_model(herit_model, 2)` and the Positive
treatment `r print_model(herit_model, 4)`. Notably, the sign of the effect is
reversed: negative for the Neutral treatment and positive for the Positive
treatment.

The negative heritability of methane oxidation rate in the Neutral selection
treatment may result from selection on alternative characters in those
populations. The strong, positive heritability in the Positive treatment ... .
It seems strange that the Methane oxidation rate was heritable between
consecutive passages (`r figures("herit", display="cite")`). Specifically, the
narrow-sense heritability, calculated as the slope of the mid-parent and
mid-offspring, was $h^2$ =  $\pm$  (t = , p = ). With an inbreeding coefficient of ,

@@

@@

@@

These experiments would further help lus understand the nature of the "map" between community structure and ecosystem function. This would not only reveal how best to predict ecosystem function in earth system models, but also demonstrate the mechanism underlying that relationship. This is akin to deducing the genetic control of trait variation in molecular biology.

@@

>>>>

---
title: Artificial ecosystem selection to deduce the mapping between microbial community structure and ecosystem function
author: "Andrew H. Morris and Brendan J. M. Bohannan"
output:
  word_document:
    reference_docx: reference.docx
bibliography: bibliography.bib
editor_options:
  chunk_output_type: inline
---

@@

```{r initialize_packages}
library(tidyverse)
library(captioner)
library(broom)
library(broman)
library(morris)
```

```{r load_data}
heritability <- read_tsv('../Output/heritability.tsv')
fluxes <- read_tsv('../Output/fluxes.tsv')

@@

figures <- captioner()

figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression. The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant `k` (i.e., `-k`) with units day^-1^. The x-axis is the passage number. The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment. The dark points are the selected jars in each passage and the light points are the non-selected jars. Regression lines aIn passage 2, the jars with the small `-k` were selected due to an error in the flux calculations.")

figures("herit", "Ordinary least squares regression of mid-offspring CH~4~ flux on mid-parent CH~4~ flux. Mid-parent is the mean of the jars selected to inoculate the next passage. Mid-offspring are the mean of all twelve jars produced in one passage. The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment.")

@@

Soil microbial communities are central to the functioning of ecosystems
[@crowther2019]. And yet, microbes have historically been excluded from
ecosystems models in favor of abiotic variables [@schimel1995]. This is in part
because attempts to establish quantitative relationships between microbial
marker genes and ecosystem functions have not been fruitful [@graham2016,
@rocca2015]. Here we use a novel approach to estimate how much of the variation
in ecosystem function can be attributed to variation in the microbial community.
We then identify microbial markers of ecosystem function. These experiments will
expand our ability to model microbial ecosystem-function relationships and allow
us to manage ecosystems for particular ecosystem functions [@fuhrman2009,
@schimel1998].

In macroorganismal communities, there appears to be a relationship between the
diversity of a community and the resilience of ecosystem functions to
environmental perturbations [@hooper2005]. Typically, diversity is represented
as the richness of a particular functional group, such as photosynthetic plants.
Similar studies in microbial communities rarely find a relationship between
diversity and function [@graham2016]. However, these studies define
diversity much more broadly by looking at the richness of the entire bacterial
and archaeal community using 16S rRNA gene variants. Studies that focus more
narrowly on a particular functional group still rarely find a significant
correlation between the abundance of a functional group and the rate of the
corresponding process [@rocca2015]. For certain functions in certain
ecosystems, these relationships can be informative. For example, the abundance
of a single methanogenic species (Candidatus ‘*Methanoflorens
stordalenmirensis*’) is correlated with the CH~4~:CO~2~ ratio in a permafrost
ecosystem [@mccalley2014]. In addition, bacterial and archaeal community composition is
correlated with methane emissions in forests and pastures of the Brazilian
Amazon [@meyer2020]. Therefore, there appears to be some indication that
microbial community structure drives variation in methane cycling in soils.

Why have we been unsuccessful at identifying relationships between
microbial community structure and ecosystem functions? Microbial communities are
extremely diverse, particularly in soil environments [@whitman1998].
Characterizing these communities using broad phylogenetic markers, such as 16S,
reveals thousands of mostly uninformative taxa which makes it difficult to
identify markers of ecosystem function. Some studies constrain this diversity by
characterizing individual protein-coding genes. However, this approach assumes
investigators already know what limits the rate of a process in a particular
ecosystem. Typically, this is assumed to be the gene that encodes the final step
in a pathway--for example, the *mcrA* gene, which encodes a subunit of the
enzyme that performs the final step in methanogenesis. However, regulation may
happen further upstream (or downstream) within the pathway and therefore the
abundance of the final gene would not correlate with the rate of the process. In
the case of methane emissions, the limitation may be in generating the
substrates for methanogenesis or may be in the consumption of methane by
methanotrophs.

To further investigate the relationship between microbial community structure
and ecosystem function, we want to determine how much of the variation in soil
methane cycling is attributable to variation in the microbial community. This
ecological question is analogous to the problem of estimating heritability in
quantitative genetics [@morris2020]. Variation in an organismal trait is
determined by the sum of genetic variation, environmental variation, and the interaction between the two [@falconer1996]. One goal of quantitative genetics is to determine the
proportion of the variation attributable to genetics. This is commonly
estimated as the narrow-sense heritability (h^2^) defined as the additive
genetic variance. This goal can be achieved through artificial selection
experiments with selection on the character of interest [@falconer1996]. We
applied artificial ecosystem selection to whole microbial communities by
selecting on the ecosystem-scale methane oxidation rate in order to estimate the
additive genetic variance of soil methane oxidation [@swenson2000].

To identify which organisms are important for regulating methane cycling, we
compared the composition of the artificially selected microbial
communities and a control set of communities without selection. Artificial
ecosystem selection has a similar effect to enrichment culturing by amplifying
the population of interest - in this case methanotrophs [@swenson2000]. This will reduce the
diversity of soil microbes and allow for greater power in detecting significant
markers of methane oxidation rate. To do this, we extracted DNA from several of
the ecosystems within the selection experiment and sequenced the 16S rRNA gene
in order to perform a differential abundance test on methane oxidation rate.

@@

## Selection experiment
We performed an artificial ecosystem selection experiment [*sensu* @swenson2000]
by passaging replicate soil microbial communities. The trait we selected on was
CH~4~ oxidation rate. Our experiment had two selection lines with twelve jars
each for a total of twenty-four jars per passage. One line was a positive
selection line where the two or three jars with the highest CH~4~ oxidation rate
were chosen to inoculate the next set of positive jars. The other line was a
neutral selection line where an equal number of jars as the positive line were
chosen at random to inoculate the next set of neutral jars. The number of jars
chosen was based on the distribution of fluxes among the positive jars, i.e., we
chose the top three jars unless only two jars had considerably greater oxidation
rates based on visual inspection of histograms. The experiment was carried out
over five passages.

The initial soil microbial community was sampled from the top 10 cm of an upland
mineral soil under a deciduous forest ecosystem near the University of Oregon
campus in Eugene, OR, USA. Incubations were performed in 500 mL mason jars with
rubber septa installed in the lids. Each jar was sterilized with 70% ethanol to
which was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL
of sterile deionized water to bring them to 60% of field capacity. The jars were
then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
headspace concentration to approximately 1000 ppm CH~4~. To create the two
treatment groups, twenty-four jars were created in an identical manner and then
randomly assigned to either the positive or neutral selection line. Jars were
flushed and respiked twice per week to maintain aerobic conditions and elevated
CH~4~ concentrations and were incubated at ambient temperature for approximately
three weeks. Methane oxidation rates were determined at the end of the
incubation period. These values were then used to determine the jars selected to
inoculate the next set of jars.

@@

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from each
jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for
a 5-point curve. Samples were immediately injected into a SRI model 8610C gas
chromatograph (Torrance, CA, USA) to determine the headspace CH~4~
concentration. Fluxes were calculated from a first-order exponential decay
function as decay constant `k` with units day^-1^. Oxidation rates are presented
as the additive inverse of `k` (i.e., `-k`) so that a more positive value
represents a greater oxidation rate. The jars selected for the positive
treatment in passage 2 had the lowest methane oxidation rate of the twelve jars
due to a calculation error in the methane oxidation rate. All other passages
correctly used the top three jars.

@@

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at -80$^\circ$C. Soil DNA was extracted from 0.25
g soil. Negative controls were extracted from autoclaved potting mix and
DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
(QIAGEN, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, USA). In order to estimate the diversity and relative abundance of
the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
region of the 16S rRNA gene using the using the 515F − 806R primer combination
[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
$\mu$l of DNA template. Reaction conditions were: 98$^\circ$C for 30 s
(initialization), 35 cycles of 98$^\circ$C for 10 s (denaturation), 61$^\circ$C
for 20 s (annealing), and 72$^\circ$C for 20 s (extension), and 72$^\circ$C for
2 m (final extension). Amplicons were purified twice using 0.8x ratio Mag-Bind
RxnPure Plus isolation beads (Omega Bio-Tek, USA). Sequencing libraries were
prepared using a dual-indexing approach (Fadrosh et al., 2014; Kozich et al.,
2013). Amplicon concentrations were quantified using Qubit (Thermo Fisher
Scientific Technologies, USA) multiplexed at equimolar concentration. Sequencing
was performed at the University of Oregon Genomics Core Facility on the Illumina
NovaSeq 6000 with paired-end 150 bp reads.

@@

Bioinformatics processing was performed in `R` (R Core Team 2021). Demultiplexed
sequencing reads were denoised using `DADA2` [@callahan2016]. Taxonomic
assignment for the 16S reads was performed using the RDP naive Bayesian
classifier [@wang2007].

@@

To test whether there was a significant change in CH~4~ oxidation rate (k) as a
response to selection, we tested a difference in slopes between the positive
and neutral selection lines. The CH~4~ oxidation rates were strongly right
skewed with most values close to zero and few large, positive values. This
resulted in residuals that did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed to better meet the assumptions of a linear model and to make
figures easier to interpret. First, we tested the effect of treatment by
fitting two nested models with and without treatment using `lm` in `R`. The
model formulation for the full model was `flux ~ passage * treatment` and the
model formulation for the reduced model was `flux ~ passage`. We compared these
models using the likelihood ratio test with the `anova` function in `R`. We
then fit the full model to estimate the slope of the positive line, which
represents the change in CH~4~ oxidation rate per passage as a response to
selection.

@@

To estimate the proportion of variance in ecosystem function due to variation in
the microbial community, we estimated narrow-sense heritability ($h^2$) as the
regression of offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the offspring was the mean for all jars
produced by those parents. First, we tested if there was an effect of treatment
on the heritability estiamte. The model formulation for the full model was
`offspring ~ parent * treatment` and the model formulation for the reduced model
was `offspring ~ parent`. We compared these models using the likelihood ratio
test with the `anova` function `R`. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

### Community response to selection

Diversity estimates and tests were performed using `DivNet` [@willis2020]. We
tested a difference in Shannon diversity between the Positive and Neutral
treatment and between the beginning and the end of the experiment with model
formulation `diversity ~ treatment * passage`. We also tested a difference in
beta-diversity as both Bray-Curtis distance and Jaccard distance between the
Positive and Neutral treatment and the beginning and end of the experiment with
model formulation `diversity ~ treatment * passage`.

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance using a beta-binomial model [@martin2020]. We compared
taxa in passage 5 between the positive and neutral treatment to identify taxa
that were enriched or depleted as a response to selection. We also tested
differential abundance with differences in CH~4~ oxidation rate to identify taxa
correlated with CH~4~ flux.

@@

Test if random forest or gradient boosting can predict whether this is the positive line or the neutral line based solely on the microbial community?

@@

## Methane oxidation

@@

We observed a response to selection on soil methane oxidation rate at the whole
ecosystem level (`r figures("response", display="cite")`). In addition, the
response to selection varied with treatment 
`r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
treatment had a mean methane oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in methane oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`% increase in methane
oxidation rate per passage `r print_lm(response_model, 4)`.

@@

We estimated heritability as the regression of mid-offspring on mid-parent.
Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
Neutral treatment `r print_lm(herit_model, 2)`. Notably, the sign of the effect
was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.

@@

## Community Response

Alpha diversity
- treatment
- flux

Beta diversity
- treatment
- flux

Differential abundance
- treatment
- flux

@@

We observed a response to selection on soil CH~4~ oxidation rate of the whole soil microbial community. This type of response has been observed for other functions, such as chloroaniline degradation in water [@swenson2000], but is the first, to our knowledge, of selection being performed on soil biogeochemical cycling.

A previous study near our sample site observed zero CH~4~ emissions from soils (cite laurel's paper). This suggests that little or no methanogenesis is happening in soils in this region and so the soil microbial communities are not "primed" to consume CH~4~. With no soil methanogenesis, the only type of methanotrophy happening would be high-affinity CH~4~ oxidation or consumption of atmospheric CH~4~, which is on the order of *some number* ppm CH~4~. By contrast, our incubations were at ~1000 ppm CH~4~. This is disparity is reflected in our results in that within the first passage we generated, only four of the 24 jars consumed more than 10% of the headspace CH~4~ within 48 hours and the biggest consumer only consumed 60% By contrast, in the final generation there were eight jars that consumed more than 10% and the top jar consumed over 95% of the headspace CH~4~.

If this experiment had been run for longer, we might be expected to see crashes in methane oxidation as the process of random sampling at each generation

Ecosystem heritability sounds confusing because it is difficult to imagine heritability without discrete generations. And how could a soil have "generations" (in the absence of coalescence events (cite Williams and Lenton))? However, this definition of heritability is conflated with the colloquial use of the term as referring to *inheritance*. But looking at the definition of narrow-sense heritability, it is simply the additive genetic variation associated with a trait [@falconer1996]. In this case, additive genetic effect would refer to the identity and relative abundance of microbial taxa in the metagenome. This notably excludes any interactions among taxa and interactions with the environment (i.e., GxE effects or context-dependent effects).

The heritability we observed was quite large at h^2^ = `r myround(tidy(herit_model)[4, 2], 2)`. However, given the small sample size and and wide confidence intervals (95% CI: `r myround(confint(herit_model)[4, ], 2)`) it is likely that a larger study could more precisely estimate the true heritability. Given these data, though, it would seem that genetic variation within the soil microbial community has a very strong effect on soil CH~4~ oxidation rates independent of any environmental variation. Granted, we intentionally eliminated much of the environmental variation that would be present in natural ecosystems.

The negative heritability of CH~4~ oxidation rate in the Neutral selection
treatment may result from selection on alternative characters in those
populations.

@@

These experiments would further help us understand the nature of the "map" between community structure and ecosystem function. This would not only reveal how best to predict ecosystem function in earth system models, but also demonstrate the mechanism underlying that relationship. This is akin to deducing the genetic control of trait variation in molecular biology.

@@

# Supplement

First, we found an effect of treatment by comparing the full model to the reduced model `r print_lrt(compare_response_mods)`.

# Junk

### 2021-08-17 R/01-process-flux.Rmd
update manuscript and analyses and figures plus add citations

@@

@@

#flux_data %>% 
#  ggplot(aes(days, ppm)) + 
#  facet_grid(flux_date + treat ~ rep, scales = "free") +
#  geom_point() + 
#  stat_smooth(method = 'lm', formula = y ~ x)

@@

fluxes <-
  nested %>% 
  mutate(tidy = map(model, tidy)) %>% 
  unnest(tidy) %>% 
  filter(term == 'days') %>% 
  group_by(flux_date) %>% 
  arrange(estimate, .by_group = TRUE) %>% 
  select(-data, -model, -term) %>% 
  ungroup() %>% 
  mutate(passage = as.numeric(flux_date)) %>% 
  mutate(ch4 = estimate,
         rank_ch4 = rank(ch4),
         log_ch4 = log10(ch4 + abs(min(ch4)) + 0.001)) %>% 
  filter(passage != 6) %>% 
  select(-std.error, -statistic, -p.value, -estimate)
```

>>>>

```{r remove_p6}
conc_data <- 
  conc_data %>% 
  filter(flux_date != '2020-07-27')
time_data <- 
  time_data %>% 
  filter(flux_date != '2020-07-27')

@@

```{r plot_fractions}

conc_data %>% 
  select(flux_date, jar, t1, t4) %>% 
  mutate(frac_cons = t4/t1) %>% 
  mutate(treat = substring(jar,1,1)) %>% 
  ggplot(aes(x = flux_date, y = frac_cons, color = treat)) +
  geom_jitter(width = 0.2) +
  geom_hline(yintercept = 0.90) +
  geom_hline(yintercept = 0.05)

@@

@@

fluxes <-
  nested %>% 
  mutate(tidy = map(model, tidy)) %>% 
  unnest(tidy) %>% 
  filter(term == 'days') %>% 
  group_by(flux_date) %>% 
  arrange(estimate, .by_group = TRUE) %>% 
  select(-data, -model, -term) %>% 
  ungroup() %>% 
  mutate(passage = as.numeric(flux_date)) %>% 
  mutate(ch4 = estimate,
         rank_ch4 = rank(ch4),
         log_ch4 = log10(ch4 + abs(min(ch4)) + 0.001)) %>% 
  select(-std.error, -statistic, -p.value, -estimate)
```

```{r plot_fluxes, fig.height=50}
flux_data %>%
 ggplot(aes(days, ppm)) +
 facet_grid(passage + treat ~ rep, scales = "free") +
 geom_point() +
 stat_smooth(method = 'lm', formula = y ~ x)
```

### 2021-08-17 R/03-response.Rmd
update manuscript and analyses and figures plus add citations

```{r compare_models}
# Compare interaction to no interaction
compare_mods <- anova(response_model, no_treat_model)
kable(tidy(compare_mods))
```

@@

```{r save_model}
save(response_model, per_slope, file = '../Output/response_model.Rdata')
```

>>>>

```{r compare_models}
# Compare interaction to no interaction
compare_response_mods <- anova(no_treat_model, response_model, test = 'LRT')
kable(tidy(compare_response_mods))
```

@@

```{r save_model}
save(response_model, compare_response_mods, per_slope, file = '../Output/response_model.Rdata')
```

### 2021-08-17 R/04-heritability.Rmd
update manuscript and analyses and figures plus add citations

Here, I fit two models. The first is simply offspring flux on selected parent flux, ignoring treatment. There is no correlation `r print_model(fit_no_treat, 2)`. Next, I added treatment and the interaction between treatment and parental flux. Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_model(herit_model, 2)` and the Positive treatment `r print_model(herit_model, 4)`. Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.

>>>>

Here, I fit two models. The first is simply offspring flux on selected parent flux, ignoring treatment. There is no correlation `r print_lm(fit_no_treat, 2)`. Next, I added treatment and the interaction between treatment and parental flux. Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_lm(herit_model, 2)` and the Positive treatment `r print_lm(herit_model, 4)`. Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.
